---
title: "Reproducible Research 1"
author: "ecvizie"
date: "March 20, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Assignment #1 - Reproducible Research
## Loading and preprocessing the data
Data taken from <https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip> on 03/20/2016. We'll use a set without NAs to start so I don't get a headache.
```{r readdata}
c <- read.csv('activity.csv')
d <- na.omit(c) #Create a copy with omitted NAs
d$date <- as.Date(d$date, '%Y-%m-%d') #Convert dates            
agg <- aggregate(d$steps, list(d$date), FUN=sum) #Aggregate and return total steps per day
hist(agg$x, main="Total # of Steps Taken Per Day", xlab="Steps Taken per Day")
```

##Calculate Summary Data for the Daily Data Set
```{r calcAvgMed}
summary(agg$x)
```

##Time series plot of the average number of steps taken
```{r timeseries1}
#Aggregate by interval across all days
agginterval <- aggregate(d$steps, list(d$interval), FUN=mean)
#Time series plot by interval
plot(x ~ Group.1, agginterval, type = "l", 
     main="Avg Steps Per Interval", xlab='Interval', ylab="Avg # of Steps")
```

##The 5-minute interval that, on average, contains the maximum number of steps
```{r maxinterval}
maxint<- agginterval[agginterval$x==max(agginterval$x),]
maxint$Group.1
```

##Code to describe and show a strategy for imputing missing data
We will use the average for that interval to fill the missing data. We accomplish this by merging datasets based on the interval period and then setting the NAs to the constant value.
```{r missingdata}
#How many NAs do we have?
length(c[is.na(c)])
#use interval average to fill data
merged <- merge(agginterval, c, by.x="Group.1", by.y = "interval")
merged[is.na(merged$steps)==TRUE,]$steps <- merged[is.na(merged$steps)==TRUE,]$x
merged$date <- as.Date(merged$date, '%Y-%m-%d')
merged$dow <- weekdays(merged$date, abbreviate = TRUE)
merged$wknd <- (merged$dow %in% c("Sat", "Sun"))
aggm <- aggregate(merged$steps, list(merged$date), FUN=sum)
summary(aggm$x)
#Histogram of Filled Data
hist(aggm$x, main="Total # of Steps Taken Per Day", xlab="Steps Taken per Day")
```

##Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
Group by the weekend feature and then by interval feature producing an average number of steps.
```{r weekenddaycompare}
aggminterval <- aggregate(merged$steps, list(merged$wknd, merged$Group.1), FUN=mean)

par(mfrow=c(2,1))
plot(x ~ Group.2, aggminterval[aggminterval$Group.1==TRUE,], type = "l",
     main="Steps taken per day (Weekend)", xlab="", ylab="# of Steps")
plot(x ~ Group.2, aggminterval[aggminterval$Group.1==FALSE,], type = "l",
     main="Steps taken per day (Weekday)", xlab="Interval",ylab="# of Steps")

```
